---
title: "Regression Analysis — Effect Modification by Baseline WBC"
author: "Dan Rzewnicki"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  warning = FALSE,
  message = FALSE
)
```

## 1. Load packages

```{r packages}
required_packages <- c(
  "tidyverse", "here", "gtsummary", "broom",
  "lmtest", "MASS", "knitr"
)

installed <- rownames(installed.packages())
to_install <- setdiff(required_packages, installed)
if (length(to_install) > 0) install.packages(to_install)

library(tidyverse)
library(here)
library(gtsummary)
library(broom)
library(lmtest)
library(MASS)
library(knitr)
```

---

## 2. Load data

```{r load-data}
df_analysis <- readRDS(here("data", "analysis_dataset.rds"))

cat("Dimensions:", nrow(df_analysis), "rows ×", ncol(df_analysis), "columns\n")
```

---

## 3. Model functions

These two functions are defined once and reused across all outcomes and sensitivity
analyses below, ensuring every model uses identical covariate sets.

```{r model-functions}
# Covariates used in every model — update here if the covariate set changes
covariates <- c(
  "trt_group",
  "baseline_wbc",
  "age",
  "sex",
  "baseline_sofa",
  "baseline_vent",
  "baseline_anyvasopressor",
  "source_of_infection",
  "location",
  "preill_creat",
  "priorhd"
)

# Interaction model: includes baseline_wbc * trt_group term
interaction_model <- function(outcome_var, data) {
  f <- as.formula(
    paste(outcome_var, "~",
          paste(covariates, collapse = " + "),
          "+ baseline_wbc:trt_group")
  )
  glm(f, data = data, family = "binomial")
}

# Main effects model: no interaction term
no_interaction_model <- function(outcome_var, data) {
  f <- as.formula(
    paste(outcome_var, "~", paste(covariates, collapse = " + "))
  )
  glm(f, data = data, family = "binomial")
}

# Helper: extract OR (95% CI) and p-value from a glm object
extract_results <- function(model) {
  mod_summary <- summary(model)
  results <- cbind(
    OR     = exp(coef(model)),
    exp(confint(model)),
    p      = mod_summary$coefficients[, "Pr(>|z|)"]
  ) %>%
    as.data.frame() %>%
    rownames_to_column("term") %>%
    rename(ci_low = `2.5 %`, ci_high = `97.5 %`) %>%
    mutate(across(c(OR, ci_low, ci_high), ~ round(.x, 3)),
           p = round(p, 4))
  results
}

# Helper: format results as a clean kable table
results_table <- function(results, caption = NULL) {
  knitr::kable(
    results,
    caption   = caption,
    col.names = c("Term", "OR", "95% CI low", "95% CI high", "p-value"),
    align     = c("l", "r", "r", "r", "r")
  )
}
```

---

## 4. Primary outcome — 28-day in-hospital mortality

### 4a. Fit models

```{r mortality-models}
mortality_int  <- interaction_model("inhosp_mortality28", analysis_df)
mortality_main <- no_interaction_model("inhosp_mortality28", analysis_df)
```

### 4b. Likelihood ratio test — interaction vs. main effects

```{r mortality-lrt}
lrtest(mortality_int, mortality_main)
```

### 4c. Interaction model results

```{r mortality-int-results}
mortality_int_results <- extract_results(mortality_int)
results_table(mortality_int_results,
              caption = "Table 2a. 28-day mortality — interaction model")
```

### 4d. Main effects model results

```{r mortality-main-results}
mortality_main_results <- extract_results(mortality_main)
results_table(mortality_main_results,
              caption = "Table 2b. 28-day mortality — main effects model")
```

### 4e. Formatted gtsummary table (for manuscript)

```{r mortality-gtsummary}
mortality_int %>%
  tbl_regression(
    exponentiate = TRUE,
    label = list(
      trt_group               ~ "Treatment group",
      baseline_wbc            ~ "Baseline WBC",
      age                     ~ "Age",
      sex                     ~ "Sex",
      baseline_sofa           ~ "Baseline SOFA score",
      baseline_vent           ~ "Mechanical ventilation",
      baseline_anyvasopressor ~ "Vasopressor use",
      source_of_infection     ~ "Source of infection",
      location                ~ "Location",
      preill_creat            ~ "Pre-illness creatinine",
      priorhd                 ~ "Prior heart disease"
    )
  ) %>%
  bold_p(t = 0.05) %>%
  modify_caption("Table 2: 28-day in-hospital mortality (interaction model)")
```

---

## 5. Secondary outcome — MAKE-14

### 5a. Fit models

```{r make14-models}
make14_int  <- interaction_model("MAKE14", analysis_df)
make14_main <- no_interaction_model("MAKE14", analysis_df)
```

### 5b. Likelihood ratio test

```{r make14-lrt}
lrtest(make14_int, make14_main)
```

### 5c. Results

```{r make14-results}
results_table(extract_results(make14_int),
              caption = "Table 3a. MAKE-14 — interaction model")

results_table(extract_results(make14_main),
              caption = "Table 3b. MAKE-14 — main effects model")
```

---

## 6. Secondary outcome — Incident AKI (ordinal)

Incident AKI is modeled using ordinal logistic regression (`MASS::polr`) given its
ordered categorical nature. p-values are calculated from z-scores since polr does
not return them directly.

### 6a. Fit models

```{r aki-models}
aki_int <- MASS::polr(
  incident_aki_with_onstudy_14.factor ~
    trt_group + baseline_wbc + baseline_wbc:trt_group +
    age + sex + baseline_sofa + baseline_vent +
    baseline_anyvasopressor + preill_creat + priorhd +
    source_of_infection + location,
  data = analysis_df,
  Hess = TRUE
)

aki_main <- MASS::polr(
  incident_aki_with_onstudy_14.factor ~
    trt_group + baseline_wbc +
    age + sex + baseline_sofa + baseline_vent +
    baseline_anyvasopressor + preill_creat + priorhd +
    source_of_infection + location,
  data = analysis_df,
  Hess = TRUE
)
```

### 6b. Likelihood ratio test

```{r aki-lrt}
anova(aki_int, aki_main, test = "Chisq")
```

### 6c. Interaction model results with p-values

```{r aki-int-results}
aki_coefs <- coef(summary(aki_int))
z_values  <- aki_coefs[, "Value"] / aki_coefs[, "Std. Error"]
p_values  <- 2 * pnorm(abs(z_values), lower.tail = FALSE)

aki_int_results <- data.frame(
  term    = rownames(aki_coefs),
  OR      = round(exp(aki_coefs[, "Value"]), 3),
  ci_low  = round(exp(confint(aki_int)[, 1]), 3),
  ci_high = round(exp(confint(aki_int)[, 2]), 3),
  p       = round(p_values, 4)
)

knitr::kable(
  aki_int_results,
  caption   = "Table 4. Incident AKI — ordinal logistic regression (interaction model)",
  col.names = c("Term", "OR", "95% CI low", "95% CI high", "p-value")
)
```

---

## 7. Secondary outcome — Coma or delirium (comagdelc14)

### 7a. Fit models

```{r coma-models}
coma_int  <- interaction_model("comagdelc14", analysis_df)
coma_main <- no_interaction_model("comagdelc14", analysis_df)
```

### 7b. Likelihood ratio test

```{r coma-lrt}
lrtest(coma_int, coma_main)
```

### 7c. Results

```{r coma-results}
results_table(extract_results(coma_int),
              caption = "Table 5a. Coma/delirium — interaction model")

results_table(extract_results(coma_main),
              caption = "Table 5b. Coma/delirium — main effects model")
```

---

## 8. Sensitivity analyses

All sensitivity analyses use 28-day in-hospital mortality as the outcome and follow
the same interaction vs. main effects structure as the primary analysis.

```{r sensitivity-function}
# Runs both models, LRT, and prints tables for a given data subset
run_sensitivity <- function(data, label) {

  mod_int  <- interaction_model("inhosp_mortality28", data)
  mod_main <- no_interaction_model("inhosp_mortality28", data)

  list(
    lrt          = lrtest(mod_int, mod_main),
    int_results  = extract_results(mod_int),
    main_results = extract_results(mod_main)
  )
}
```

### 8a. Sepsis + vancomycin recipients only

```{r sens-sepsis-vanc}
sepsis_vanc      <- analysis_df %>%
  filter(received_vancomycin_14 == 1 & sepsis == "Yes")
sens_sv          <- run_sensitivity(sepsis_vanc, "Sensitivity: sepsis + vancomycin")

print(sens_sv$lrt)
results_table(sens_sv$int_results,  "Sensitivity 1a. Sepsis + vancomycin — interaction")
results_table(sens_sv$main_results, "Sensitivity 1b. Sepsis + vancomycin — main effects")
```

### 8b. Vancomycin recipients only

```{r sens-vanc}
vanc        <- analysis_df %>% filter(received_vancomycin_14 == 1)
sens_vanc   <- run_sensitivity(vanc, "Sensitivity: vancomycin only")

print(sens_vanc$lrt)
results_table(sens_vanc$int_results,  "Sensitivity 2a. Vancomycin only — interaction")
results_table(sens_vanc$main_results, "Sensitivity 2b. Vancomycin only — main effects")
```

### 8c. Baseline sepsis only

```{r sens-sepsis}
sepsis       <- analysis_df %>% filter(sepsis == "Yes")
sens_sepsis  <- run_sensitivity(sepsis, "Sensitivity: sepsis only")

print(sens_sepsis$lrt)
results_table(sens_sepsis$int_results,  "Sensitivity 3a. Sepsis only — interaction")
results_table(sens_sepsis$main_results, "Sensitivity 3b. Sepsis only — main effects")
```

### 8d. No sepsis at baseline

```{r sens-no-sepsis}
no_sepsis       <- analysis_df %>% filter(sepsis == "No")
sens_no_sepsis  <- run_sensitivity(no_sepsis, "Sensitivity: no sepsis")

print(sens_no_sepsis$lrt)
results_table(sens_no_sepsis$int_results,  "Sensitivity 4a. No sepsis — interaction")
results_table(sens_no_sepsis$main_results, "Sensitivity 4b. No sepsis — main effects")
```

### 8e. Excluding intra-abdominal infections

```{r sens-no-abdominal}
no_abdominal <- analysis_df %>%
  filter(source_of_infection != "Intra-abdominal") %>%
  mutate(source_of_infection = droplevels(source_of_infection))
sens_no_abd  <- run_sensitivity(no_abdominal, "Sensitivity: exclude intra-abdominal")

print(sens_no_abd$lrt)
results_table(sens_no_abd$int_results,  "Sensitivity 5a. No intra-abdominal — interaction")
results_table(sens_no_abd$main_results, "Sensitivity 5b. No intra-abdominal — main effects")
```

### 8f. Excluding neutropenic patients

```{r sens-no-neutropenia}

no_neutropenia <- analysis_df %>%
  filter(baseline_neutroenia == "No") 
sens_no_neutropenia  <- run_sensitivity(no_neutropenia, "Sensitivity: exclude neutropenic patients")

print(sens_no_neutropenia$lrt)
results_table(sens_no_neutropenia$int_results,  "Sensitivity 5a. No Neutropenia — interaction")
results_table(sens_no_neutropenia$main_results, "Sensitivity 5b. No Neutropenia — main effects")

```

### 8g. Sensitivity — adding neutropenia as a covariate

```{r sens-neutropenia}
# Extend the standard covariate set for this analysis only
covariates_neutropenia <- c(covariates, "neutropenia")

# Define models inline — does not affect the global covariates vector
neut_int <- glm(
  as.formula(
    paste("inhosp_mortality28 ~",
          paste(covariates_neutropenia, collapse = " + "),
          "+ baseline_wbc:trt_group")
  ),
  data   = analysis_df,
  family = "binomial"
)

neut_main <- glm(
  as.formula(
    paste("inhosp_mortality28 ~",
          paste(covariates_neutropenia, collapse = " + "))
  ),
  data   = smart_abx,
  family = "binomial"
)

lrtest(neut_int, neut_main)

results_table(extract_results(neut_int),
              caption = "Sensitivity 6a. Neutropenia added — interaction model")
results_table(extract_results(neut_main),
              caption = "Sensitivity 6b. Neutropenia added — main effects model")
```


##9. Analyses by WBC Quartile


###9a. 28-day mortality using WBC Quartile

```{r quartile covariates}

# Covariates used in every model — update here if the covariate set changes
quartile_covariates <- c(
  "trt_group",
  "wbc_quartile",
  "age",
  "sex",
  "baseline_sofa",
  "baseline_vent",
  "baseline_anyvasopressor",
  "source_of_infection",
  "location",
  "preill_creat",
  "priorhd"
)
```


```{r quartile functions}

# Interaction model: includes baseline_wbc * trt_group term
quartile_interaction_model <- function(outcome_var, data) {
  f <- as.formula(
    paste(outcome_var, "~",
          paste(covariates, collapse = " + "),
          "+ wbc_quartile:trt_group")
  )
  glm(f, data = data, family = "binomial")
}

# Main effects model: no interaction term
quartile_no_interaction_model <- function(outcome_var, data) {
  f <- as.formula(
    paste(outcome_var, "~", paste(quartile_covariates, collapse = " + "))
  )
  glm(f, data = data, family = "binomial")
}

```


```{r mortality-models-quartile}
mortality_int_quartile  <- quartile_interaction_model("inhosp_mortality28", analysis_df)
mortality_main_quartile <- quartile_no_interaction_model("inhosp_mortality28", analysis_df)
```



```{r quartile-mortality-lrt}
lrtest(mortality_int_quartile, mortality_main_quartile)
```

### 4c. Interaction model results

```{r mortality-int-results}
mortality_int_results_quartile <- extract_results(mortality_int_quartile)
results_table(mortality_int_results_quartile,
              caption = "Table 2a. 28-day mortality — interaction model by WBC Quartile")
```


```{r mortality-main-results-quartile}
mortality_main_results_quartile <- extract_results(mortality_main_quartile)
results_table(mortality_main_results_quartile,
              caption = "Table 2b. 28-day mortality — main effects model by WBC Quartile")
```

## 10. Export all results to CSV

```{r export}
dir.create(here("output"), showWarnings = FALSE)

# Primary outcome
write.csv(mortality_int_results,
          here("output", "results_mortality_interaction.csv"),  row.names = FALSE)
write.csv(mortality_main_results,
          here("output", "results_mortality_main.csv"),         row.names = FALSE)

# Secondary outcomes
write.csv(extract_results(make14_int),
          here("output", "results_make14_interaction.csv"),     row.names = FALSE)
write.csv(extract_results(coma_int),
          here("output", "results_coma_interaction.csv"),       row.names = FALSE)
write.csv(aki_int_results,
          here("output", "results_aki_interaction.csv"),        row.names = FALSE)

# Sensitivity analyses
write.csv(sens_sv$int_results,
          here("output", "results_sens_sepsis_vanc.csv"),       row.names = FALSE)
write.csv(sens_vanc$int_results,
          here("output", "results_sens_vanc.csv"),              row.names = FALSE)
write.csv(sens_sepsis$int_results,
          here("output", "results_sens_sepsis.csv"),            row.names = FALSE)
write.csv(sens_no_sepsis$int_results,
          here("output", "results_sens_no_sepsis.csv"),         row.names = FALSE)
write.csv(sens_no_abd$int_results,
          here("output", "results_sens_no_abdominal.csv"),      row.names = FALSE)
write.csv(sens_no_neutropenia$int_results,
          here("output", "results_sens_no_neutropenia.csv"))
write.csv(extract_results(neut_int),
          here("output", "results_sens_neutropenia.csv"), row.names = FALSE)

#analyses by quartile
write.csv(extract_results(mortality_int_quartile),
          here("output", "results_mortality_quartile.csv"), row.names = FALSE)

cat("All results exported to output/\n")
```

---

## 10. Session info

```{r session-info}
sessionInfo()
```
