---
title: "Manuscript Figures"
author: "Dan Rzewnicki"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width  = 7,
  fig.height = 8,
  dpi        = 300
)
```

## Load packages

```{r packages}
required_packages <- c("tidyverse", "here", "broom", "purrr", "patchwork", "knitr")

installed <- rownames(installed.packages())
to_install <- setdiff(required_packages, installed)
if (length(to_install) > 0) install.packages(to_install)

library(tidyverse)
library(here)
library(broom)
library(purrr)
library(patchwork)
```

## Load data

```{r load-data}
df_analysis <- readRDS(here("data", "analysis_dataset.rds"))
```

---

## Shared theme

```{r theme}
# Apply once here; all figures below inherit it automatically
theme_manuscript <- theme_bw(base_size = 12) +
  theme(
    panel.grid.minor  = element_blank(),
    strip.background  = element_rect(fill = "grey92", color = NA),
    legend.position   = "bottom",
    plot.title        = element_text(face = "bold", size = 13),
    plot.subtitle     = element_text(size = 11, color = "grey40"),
    axis.title        = element_text(size = 11),
    axis.text         = element_text(size = 10)
  )

theme_set(theme_manuscript)
```

## Shared model setup

The covariate set and quartile labels are defined once and used for all four outcomes.

```{r shared-setup}
# Covariates used in every quartile-stratified model
quartile_covariates <- c(
  "trt_group", "age", "sex", "baseline_sofa",
  "baseline_vent", "baseline_anyvasopressor",
  "source_of_infection", "preill_creat", "priorhd"
)

# Quartile labels — update cutpoints to match your data
quartile_labels <- c("<7", "7-11", "11-16", ">16")
```

---

## Helper function — fit models by quartile

This function takes an outcome variable name and returns a tidy data frame of
odds ratios for `trt_group` within each WBC quartile. Using `tidy(conf.int = TRUE)`
avoids the list-column alignment issue that can occur when unnesting `confint()`
output separately.

```{r or-function}
get_quartile_ORs <- function(outcome_var, data, covariate_set, q_labels) {

  f <- as.formula(
    paste(outcome_var, "~", paste(covariate_set, collapse = " + "))
  )

  data %>%
    mutate(wbc_quartile = factor(wbc_quartile,
                                 levels = 1:4,
                                 labels = q_labels)) %>%
    group_by(wbc_quartile) %>%
    nest() %>%
    mutate(
      fit    = map(data, ~ glm(f, data = .x, family = "binomial")),
      output = map(fit,  ~ tidy(.x, conf.int = TRUE, exponentiate = TRUE))
    ) %>%
    unnest(output) %>%
    filter(term == "trt_groupzosyn") %>%
    select(wbc_quartile,
           OR     = estimate,
           LL     = conf.low,
           UL     = conf.high,
           p      = p.value) %>%
    mutate(across(c(OR, LL, UL), ~ round(.x, 3)),
           p = round(p, 4))
}
```

---

## Fit models for each outcome

```{r fit-all}
or_mortality <- get_quartile_ORs("inhosp_mortality28", df_analysis,
                                  quartile_covariates, quartile_labels)

or_make14    <- get_quartile_ORs("MAKE14",             df_analysis,
                                  quartile_covariates, quartile_labels)

or_coma      <- get_quartile_ORs("comagdelc14",        df_analysis,
                                  quartile_covariates, quartile_labels)

or_aki       <- get_quartile_ORs("incident_aki_with_onstudy_14.factor",
                                  df_analysis, quartile_covariates, quartile_labels)
```

---

## Figure — Combined forest plot

```{r forest-plot-function}
# Shared theme for all panels
theme_forest <- theme_manuscript +
  theme(
    panel.grid.major.y = element_blank(),   # no horizontal grid in forest plots
    axis.title.y       = element_blank(),   # y-axis label not needed (groups are self-labeling)
    plot.margin        = margin(4, 8, 4, 8)
  )

# Function to draw one panel of the forest plot
forest_panel <- function(data, title, x_limits = c(0.2, 5)) {

  ggplot(data, aes(x = OR, y = fct_rev(wbc_quartile))) +
    geom_vline(xintercept = 1, linetype = "dashed",
               color = "grey50", linewidth = 0.5) +
    geom_errorbarh(aes(xmin = LL, xmax = UL),
                   height = 0.2, linewidth = 0.7, color = "#2166ac") +
    geom_point(aes(color = OR < 1), size = 3.5) +
    scale_color_manual(
      values = c("TRUE" = "#2166ac", "FALSE" = "#d73027"),
      guide  = "none"
    ) +
    scale_x_log10(limits = x_limits,
                  breaks  = c(0.25, 0.5, 1, 2, 4),
                  labels  = c("0.25", "0.5", "1.0", "2.0", "4.0")) +
    # Add OR (95% CI) as text labels to the right of each point
    geom_text(
      aes(x     = x_limits[2],
          label = sprintf("%.2f (%.2f, %.2f)", OR, LL, UL)),
      hjust  = 1,
      size   = 3,
      color  = "grey30"
    ) +
    labs(title = title,
         x     = "Odds ratio (log scale)") +
    theme_forest
}
```

```{r combined-forest, fig.height=10}
p_mortality <- forest_panel(or_mortality, "28-day in-hospital mortality")
p_make14    <- forest_panel(or_make14,    "MAKE-14")
p_coma      <- forest_panel(or_coma,      "Coma or delirium (comagdelc14)")
p_aki       <- forest_panel(or_aki,       "Incident AKI")

# Stack all four panels using patchwork
# collect_axes = TRUE shares the x-axis label across panels
combined_forest <- (p_mortality / p_make14 / p_coma / p_aki) +
  plot_annotation(
    title    = "Figure X. Odds of outcomes by WBC quartile (Zosyn vs. comparator)",
    subtitle = "Points are odds ratios; horizontal lines are 95% confidence intervals.\nBlue = OR < 1 (favors Zosyn); red = OR > 1 (favors comparator).",
    theme    = theme(
      plot.title    = element_text(face = "bold", size = 13),
      plot.subtitle = element_text(size = 10, color = "grey40")
    )
  ) &
  theme(axis.title.x = element_text(size = 10))

combined_forest
```

---

## Save figure

```{r save-figure}
dir.create(here("output"), showWarnings = FALSE)

ggsave(here("output", "fig_quartile_forest_combined.pdf"),
       combined_forest, width = 7, height = 10, dpi = 300)

ggsave(here("output", "fig_quartile_forest_combined.tiff"),
       combined_forest, width = 7, height = 10, dpi = 300)

cat("Combined forest plot saved to output/\n")
```

---

## Export underlying data

```{r export-or-data}
# Combine all four outcome OR tables and export as one CSV
# Useful for supplementary tables or editorial review
all_or_data <- bind_rows(
  or_mortality %>% mutate(outcome = "28-day mortality"),
  or_make14    %>% mutate(outcome = "MAKE-14"),
  or_coma      %>% mutate(outcome = "Coma or delirium"),
  or_aki       %>% mutate(outcome = "Incident AKI")
) %>%
  select(outcome, wbc_quartile, OR, LL, UL, p)

write.csv(all_or_data,
          here("output", "results_quartile_forest_data.csv"),
          row.names = FALSE)

knitr::kable(
  all_or_data,
  caption   = "Odds ratios by WBC quartile for all outcomes",
  col.names = c("Outcome", "WBC Quartile", "OR", "95% CI low", "95% CI high", "p-value")
)
```

---

## Figure — Clinical cutoff forest plot (mortality only)

```{r forest with different WBC cutoff}

# Get ORs using the same helper function from the quartile section,
# but grouping by wbc_clinical instead of wbc_quartile
clinical_OR_data <- df_analysis %>%
  mutate(wbc_clinical = as.factor(case_when(
    baseline_wbc <= 4 ~ "<4",
    baseline_wbc > 4 & baseline_wbc <= 10 ~ "4-10",
    baseline_wbc > 10 ~ ">10"
  ))) %>%
  group_by(wbc_clinical) %>%
  nest() %>%
  mutate(
    fit    = map(data, ~ glm(inhosp_mortality28 ~ trt_group + age + sex +
                               baseline_sofa + baseline_vent +
                               baseline_anyvasopressor + source_of_infection +
                               preill_creat + priorhd,
                             data = .x, family = "binomial")),
    output = map(fit, ~ tidy(.x, conf.int = TRUE, exponentiate = TRUE))
  ) %>%
  unnest(output) %>%
  filter(term == "trt_groupzosyn") %>%
  select(wbc_clinical,
         OR  = estimate,
         LL  = conf.low,
         UL  = conf.high,
         p   = p.value) %>%
  mutate(across(c(OR, LL, UL), ~ round(.x, 3)),
         p           = round(p, 4),
         wbc_clinical = factor(wbc_clinical))   # ensure factor for plotting

# Reuse forest_panel() — just pass wbc_clinical as the grouping variable
# forest_panel() expects a column named wbc_quartile, so rename for compatibility
clinical_forest <- clinical_OR_data %>%
  rename(wbc_quartile = wbc_clinical) %>%
  forest_panel(
    data      = .,
    title     = "28-day in-hospital mortality — clinical WBC cutoffs",
    x_limits  = c(0.2, 5)
  ) +
  labs(
    x        = "Odds ratio (log scale)\nFavors piperacillin-tazobactam          Favors cefepime",
    caption  = "Points are odds ratios; horizontal lines are 95% confidence intervals.\nBlue = OR < 1 (favors piperacillin-tazobactam); red = OR > 1 (favors cefepime)."
  )

clinical_forest

# Save
ggsave(here("output", "fig_clinical_cutoff_forest.pdf"),
       clinical_forest, width = 7, height = 4, dpi = 300)

ggsave(here("output", "fig_clinical_cutoff_forest.tiff"),
       clinical_forest, width = 7, height = 4, dpi = 300)

```

--- 

## Figure — Predicted probability of 28-day mortality by baseline WBC and treatment

``` {r predicted morality figure}

### Fit model

# This is the same interaction model from 03_logistic_regression.Rmd.
# Re-fitted here so the figures script is self-contained and can be
# run independently without sourcing the regression script.
moda <- glm(
  inhosp_mortality28 ~ trt_group + baseline_wbc + baseline_wbc:trt_group +
    age + sex + baseline_sofa + baseline_vent + baseline_anyvasopressor +
    source_of_infection + location + preill_creat + priorhd,
  data   = df_analysis,
  family = "binomial"
)

### Build prediction grid

# Continuous WBC sequence for smooth curves
wbc_seq <- seq(1, 35, by = 0.5)

# All covariates other than WBC and treatment are held at their
# median (continuous) or most common value (categorical).
# Using median() / names(sort())[length()] rather than a hardcoded
# template row makes this robust if the dataset changes.
pred_grid <- expand.grid(
  baseline_wbc = wbc_seq,
  trt_group    = levels(df_analysis$trt_group)
) %>%
  mutate(
    age                     = median(df_analysis$age,           na.rm = TRUE),
    sex                     = names(sort(table(df_analysis$sex),          decreasing = TRUE))[1],
    baseline_sofa           = median(df_analysis$baseline_sofa, na.rm = TRUE),
    baseline_vent           = names(sort(table(df_analysis$baseline_vent), decreasing = TRUE))[1],
    baseline_anyvasopressor = names(sort(table(df_analysis$baseline_anyvasopressor), decreasing = TRUE))[1],
    source_of_infection     = names(sort(table(df_analysis$source_of_infection), decreasing = TRUE))[1],
    location                = names(sort(table(df_analysis$location),     decreasing = TRUE))[1],
    preill_creat            = median(df_analysis$preill_creat,  na.rm = TRUE),
    priorhd                 = names(sort(table(df_analysis$priorhd),      decreasing = TRUE))[1],
    # Re-apply factor levels to match the fitted model
    across(where(is.character), ~ factor(.x, levels = levels(df_analysis[[cur_column()]])))
  )

### Predict

pred_grid <- pred_grid %>%
  mutate(pred_prob = predict(moda, newdata = pred_grid, type = "response"))

### Plot

predicted_mortality_plot <- ggplot(
  pred_grid,
  aes(x = baseline_wbc, y = pred_prob, color = trt_group)
) +
  geom_line(linewidth = 1.2) +
  # Mark the three clinical cutpoints used in the forest plot
  geom_vline(xintercept = c(7, 11, 16),
             linetype = "dashed", color = "grey70", linewidth = 0.4) +
  annotate("text", x = c(7, 11, 16), y = 0,
           label = c("7", "11", "16"),
           vjust = -0.5, hjust = -0.2, size = 3, color = "grey50") +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, NA)
  ) +
  scale_color_manual(
    values = c("#0072B2", "#D55E00"),
    labels = c("zosyn"    = "Piperacillin-tazobactam",
               "cefepime" = "Cefepime")
  ) +
  labs(
    title   = "Predicted 28-day mortality by baseline WBC and treatment group",
    subtitle = paste0(
      "Other covariates held at median (continuous) or most common value (categorical).\n",
      "Dashed lines mark WBC = 7, 11, and 16."
    ),
    x     = "Baseline white blood cell count (x10\u00b3/\u03bcL)",
    y     = "Predicted probability of 28-day mortality",
    color = "Treatment group"
  ) +
  theme_manuscript +     # reuses the shared theme defined earlier in 04_figures.Rmd
  theme(
    legend.position    = "bottom",
    legend.title       = element_text(face = "bold"),
    axis.title         = element_text(face = "bold"),
    axis.text          = element_text(color = "black"),
    panel.grid.major.y = element_line(color = "gray90", linewidth = 0.3)
  )

predicted_mortality_plot

### Save

ggsave(here("output", "fig_predicted_mortality_wbc.pdf"),
       predicted_mortality_plot, width = 7, height = 5, dpi = 300)

ggsave(here("output", "fig_predicted_mortality_wbc.tiff"),
       predicted_mortality_plot, width = 7, height = 5, dpi = 300)

cat("Predicted mortality figure saved to output/\n")

```

---

## Session info

```{r session-info}
sessionInfo()
```
