---
title: "Penalized Regression — LASSO for Variable Selection"
author: "[Author Name]"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  warning = FALSE,
  message = FALSE
)
```

## Overview

This script fits a LASSO (least absolute shrinkage and selection operator) penalized
logistic regression model to identify predictors of 28-day in-hospital mortality,
including interaction terms with treatment group. The dataset is split into training
(80%) and test (20%) sets prior to model fitting. Model performance is evaluated in
the held-out test set using the area under the ROC curve (AUC) and a confusion matrix.

---

## 1. Load packages

```{r packages}
required_packages <- c(
  "tidyverse", "here", "glmnet", "caret", "pROC", "knitr"
)

installed <- rownames(installed.packages())
to_install <- setdiff(required_packages, installed)
if (length(to_install) > 0) install.packages(to_install)

library(tidyverse)
library(here)
library(glmnet)   # LASSO / ridge / elastic net
library(caret)    # createDataPartition, confusionMatrix
library(pROC)     # roc(), auc()
library(knitr)
```

---

## 2. Load data

```{r load-data}
df_analysis <- readRDS(here("data", "analysis_dataset.rds"))

cat("Dimensions:", nrow(df_analysis), "rows x", ncol(df_analysis), "columns\n")
```

---

## 3. Data preparation

### 3a. Variable cleaning

```{r cleaning}
# Variables to convert to factors
factor_vars <- c(
  "trt_group",
  "inhosp_mortality28",
  "death14_newrrt14",
  "delc14",
  "comagdelc14",
  "persistent_renal_dysfunction28",
  "newrrt28",
  "newrrt14",
  "sex",
  "CKD",
  "received_vancomycin_14",
  "received_assigned_therapy_28",
  "MAKE14",
  "baseline_vent",
  "baseline_anyvasopressor",
  "location",
  "priorhd",
  "baseline_neutropenia",
  "race_white",
  "race.eth"
)

df_analysis <- df_analysis %>%
  mutate(across(all_of(factor_vars), as.factor)) %>%
  mutate(id = as.numeric(id))
```

### 3b. Platelet — remove 12 missing

```{r platelet}
summary(df_analysis$baseline_platelet)
cat("Missing baseline_platelet:", sum(is.na(df_analysis$baseline_platelet)), "\n")

df_clean <- df_analysis %>%
  filter(!is.na(baseline_platelet))

cat("Rows after platelet filter:", nrow(df_clean), "\n")
```

### 3c. Temperature — convert Celsius recordings to Fahrenheit

Some temperatures appear to have been recorded in Celsius (values < 40). These are
converted to Fahrenheit using the standard formula. All remaining records with missing
temperature are removed.

```{r temperature}
summary(df_clean$baseline_Temp)

df_clean <- df_clean %>%
  mutate(
    clean_temp = ifelse(baseline_Temp > 40,
                        baseline_Temp,                       # already Fahrenheit
                        (baseline_Temp * (9 / 5)) + 32)     # convert from Celsius
  ) %>%
  filter(!is.na(clean_temp))

cat("Temperature range after cleaning:\n")
summary(df_clean$clean_temp)
cat("Rows after temperature filter:", nrow(df_clean), "\n")
```

### 3d. Note on arterial pH

```{r ph-note}
cat("Missing baseline_pH_art:", sum(is.na(df_clean$baseline_pH_art)), "\n")
# Arterial pH has >2200 missing values and is excluded from the model
```

---

## 4. Define predictors

Predictors are defined once here and referenced throughout. Interaction terms with
treatment group are specified separately in the model matrix formula below.

```{r predictors}
# Main effect predictors
predictors <- c(
  "age",
  "trt_group",
  "sex",
  "baseline_sofa",
  "CKD",
  "sepsis",
  "baseline_vent",
  "baseline_anyvasopressor",
  "source_of_infection",
  "location",
  "priorhd",
  "baseline_wbc",
  "baseline_neutropenia",
  "baseline_platelet",
  "baseline_hemoglobin",
  "clean_temp"
)

# Predictors that will have trt_group interaction terms
interaction_predictors <- c(
  "age",
  "baseline_wbc",
  "baseline_sofa",
  "sepsis",
  "baseline_neutropenia",
  "baseline_platelet",
  "baseline_hemoglobin",
  "clean_temp"
)

cat("Main effect predictors:", length(predictors), "\n")
cat("Interaction terms:", length(interaction_predictors), "\n")
```

---

## 5. Train / test split

The dataset is split 80% training / 20% test, stratified on baseline WBC to preserve
its distribution across both sets. A seed is set for reproducibility.

```{r split}
set.seed(123)

training_idx <- createDataPartition(
  df_clean$baseline_wbc,
  p    = 0.8,
  list = FALSE
)

train_data <- df_clean[training_idx, ]
test_data  <- df_clean[-training_idx, ]

cat("Training set:", nrow(train_data), "rows\n")
cat("Test set:    ", nrow(test_data),  "rows\n")
```

---

## 6. Build model matrices

`model.matrix()` expands categorical variables into dummy variables and creates the
interaction terms. The intercept column (column 1) is dropped since `glmnet` adds
its own.

```{r matrix-formula}
# Build interaction formula programmatically from the vectors defined in Section 4
# This avoids hard-coding the formula and makes it easy to update
interaction_terms <- paste(interaction_predictors, "trt_group", sep = ":")

lasso_formula <- as.formula(
  paste("~", paste(c(predictors, interaction_terms), collapse = " + "))
)

cat("Model formula:\n")
print(lasso_formula)
```

```{r train-matrix}
# Complete cases in training set
train_complete <- train_data %>%
  filter(complete.cases(across(all_of(c(predictors, "inhosp_mortality28")))))

cat("Complete training cases:", nrow(train_complete),
    "(removed", nrow(train_data) - nrow(train_complete), "with missing predictors)\n")

x_train <- model.matrix(lasso_formula, data = train_complete)[, -1]
y_train  <- train_complete$inhosp_mortality28
```

```{r test-matrix}
# Complete cases in test set
test_complete <- test_data %>%
  filter(complete.cases(across(all_of(c(predictors, "inhosp_mortality28")))))

cat("Complete test cases:", nrow(test_complete),
    "(removed", nrow(test_data) - nrow(test_complete), "with missing predictors)\n")

x_test <- model.matrix(lasso_formula, data = test_complete)[, -1]
y_test  <- test_complete$inhosp_mortality28
```

---

## 7. Fit LASSO model

Cross-validation is used to select the optimal penalty parameter (lambda). The
lambda that minimizes cross-validated deviance (`lambda.min`) is used for the
final model.

### 7a. Cross-validated lambda selection

```{r cv-lasso}
set.seed(123)

cv_lasso <- cv.glmnet(
  x      = x_train,
  y      = y_train,
  alpha  = 1,           # alpha = 1 is LASSO; alpha = 0 is ridge
  family = "binomial",
  nfolds = 10           # 10-fold cross-validation
)

plot(cv_lasso)
title("Cross-validation curve — log(lambda) vs. deviance", line = 2.5)

cat(sprintf("lambda.min:  %.5f  (minimizes CV error)\n", cv_lasso$lambda.min))
cat(sprintf("lambda.1se:  %.5f  (most regularized within 1 SE of minimum)\n", cv_lasso$lambda.1se))
```

### 7b. Fit final model using lambda.min

```{r final-model}
model_lasso <- glmnet(
  x      = x_train,
  y      = y_train,
  alpha  = 1,
  family = "binomial",
  lambda = cv_lasso$lambda.min
)
```

---

## 8. Model coefficients

Variables with non-zero coefficients were retained by the LASSO. A coefficient of
zero indicates the variable was penalized out of the model.

```{r coefficients}
coef_matrix <- coef(model_lasso)

coef_df <- data.frame(
  variable    = rownames(coef_matrix),
  coefficient = as.numeric(coef_matrix[, 1])
) %>%
  filter(coefficient != 0, variable != "(Intercept)") %>%
  mutate(
    OR        = round(exp(coefficient), 3),
    coefficient = round(coefficient, 4)
  ) %>%
  arrange(desc(abs(coefficient)))

knitr::kable(
  coef_df,
  caption   = "Table. Non-zero LASSO coefficients with odds ratios",
  col.names = c("Variable", "Coefficient", "Odds Ratio"),
  align     = c("l", "r", "r")
)

cat("\nVariables retained:", nrow(coef_df), "\n")
cat("Variables penalized to zero:",
    nrow(coef_matrix) - 1 - nrow(coef_df), "\n")  # subtract intercept
```

---

## 9. Model evaluation in test set

### 9a. Predicted probabilities and class predictions

```{r predictions}
test_probs <- predict(model_lasso, newx = x_test, type = "response")[, 1]

# Classify using 0.5 probability threshold
outcome_levels   <- levels(test_complete$inhosp_mortality28)
test_preds_factor <- factor(
  ifelse(test_probs > 0.5, outcome_levels[2], outcome_levels[1]),
  levels = outcome_levels
)
```

### 9b. Confusion matrix

```{r confusion-matrix}
confusionMatrix(test_preds_factor, test_complete$inhosp_mortality28)
```

### 9c. ROC curve and AUC

```{r roc-auc}
roc_obj <- roc(
  response  = test_complete$inhosp_mortality28,
  predictor = test_probs,
  quiet     = TRUE
)

cat(sprintf("AUC: %.3f (95%% CI: %.3f - %.3f)\n",
            auc(roc_obj),
            ci(roc_obj)[1],
            ci(roc_obj)[3]))

plot(roc_obj,
     main = "ROC curve — LASSO model in test set",
     col  = "#2166ac",
     lwd  = 2)
abline(a = 0, b = 1, lty = 2, col = "grey50")
```

---

## 10. Export results

```{r export}
dir.create(here("output"), showWarnings = FALSE)

write.csv(
  coef_df,
  here("output", "results_lasso_coefficients.csv"),
  row.names = FALSE
)

# Save ROC data for potential use in figures script
roc_data <- data.frame(
  specificity = roc_obj$specificities,
  sensitivity = roc_obj$sensitivities
)

write.csv(
  roc_data,
  here("output", "results_lasso_roc.csv"),
  row.names = FALSE
)

cat("Results exported to output/\n")
```

---

## 11. Session info

```{r session-info}
sessionInfo()
```
